### 2014-09-15 Pero_Transcriptome.tex
for realz

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

The filtered assemblies were annotated using the default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014. Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}. Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}. Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\

To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}. We estimated expression for the four tissues individually using default settings of the software eXpress \cite{Roberts:2012dh}. Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized. \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapter \textit{Dipodomys} rodents. The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree was constructed using the online resource Clustal-Omega, and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}. Significance was evaluated via the use of the likelihood ratio test. \\

>>>>

The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations. The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P. eremicus} dataset likely to be biological in origin. This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly. We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity. This multi-fasta file was used for all subsequent analyses, including annotation and mapping. \\

@@

The filtered assemblies were annotated using the default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014. Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}. Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}. Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\

To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \cite{Li:2013wn}. We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \cite{Roberts:2012dh}. Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized. \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapter \textit{Dipodomys} rodents. The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree was constructed using the online resource Clustal-Omega, and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}. Significance was evaluated via the use of the likelihood ratio test. \\
